---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  out.width = "100%"
)
knitr::opts_chunk$set(dpi = 300, fig.width = 7)
```

# Australia's Essential Workers

<!-- badges: start -->
<!-- badges: end -->

The goal of this repository is to identify essential workers by their ANZSIC and ANZSCO codes. Pull requests are welcome. Note that, this is by no means perfect.

# Data sources

The original data sources of the spreadsheets in 'data' folder are as stated below.

- ANZSIC: https://www.fwc.gov.au/awards-and-agreements/minimum-wages-conditions/annual-wage-reviews/previous-wage-reviews/annual-wa-0
- ANZSCO: https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1220.02013,%20Version%201.3?OpenDocument

# References

Sources of information I used to identify essential workers.

- https://www.theaustralian.com.au/world/coronavirus-the-essential-workers-wholl-keep-australia-ticking-along/news-story/cb7992b6ae722afd0838f17e61338e01
- https://www.betterteam.com/what-is-an-essential-worker-in-australia
- https://en.wikipedia.org/wiki/Essential_services
- https://manofmany.com/lifestyle/advice/what-are-non-essential-services-australia
- https://coronavirus.nt.gov.au/__data/assets/pdf_file/0006/809853/approved-remote-essential-worker-application.pdf

## Additional list 

- https://www.ama.com.au/sites/default/files/LCA_and_AMA_Essential_Services_Factsheet.pdf

# Process tables

<!-- ```{r, echo = FALSE} -->
<!-- # gt(head(anzsic_divisions, 10)) %>% -->
<!-- #   tab_header( -->
<!-- #     title = "ANZSIC 2006 divisions" -->
<!-- #   ) -->

<!-- kable(anzsic_divisions,  -->
<!--       caption = "ANZSIC 2006 divisions") -->

<!-- # gt(head(anzsic_codes, 10)) %>% -->
<!-- #   tab_header( -->
<!-- #     title = "ANZSIC 2006 codes", -->
<!-- #     subtitle = "first 10 rows only" -->
<!-- #   )  -->

<!-- kable(head(anzsic_codes, 10),  -->
<!--     caption = "ANZSIC 2006 codes, first 10 rows only") -->
<!--   # scroll_box(width = "500px", height = "200px") -->
<!-- ``` -->


<!-- ```{r, echo = FALSE} -->
<!-- kable(anzsoc_major,  -->
<!--     caption = "ANZSOC 2013 - Major group") -->

<!-- # gt(anzsoc_major) %>% -->
<!-- #   tab_header( -->
<!-- #     title = "ANZSOC 2013 - Major group" -->
<!-- #   ) -->

<!-- kable(head(anzsoc_submajor, 10),  -->
<!--     caption = "ANZSOC 2013 - Sub-major group, first 10 rows only") -->

<!-- # gt(head(anzsoc_submajor, 10)) %>% -->
<!-- #   tab_header( -->
<!-- #     title = "ANZSOC 2013 - Submajor group", -->
<!-- #     subtitle = "first 10 rows only" -->
<!-- #   ) -->

<!-- kable(head(anzsoc_minor, 10),  -->
<!--     caption = "ANZSOC 2013 - Minor group, first 10 rows only") -->

<!-- # gt(head(anzsoc_minor, 10)) %>% -->
<!-- #   tab_header( -->
<!-- #     title = "ANZSOC 2013 - Minor group", -->
<!-- #     subtitle = "first 10 rows only" -->
<!-- #   ) -->

<!-- kable(head(anzsoc_unit, 10),  -->
<!--     caption = "ANZSOC 2013 - Unit group, first 10 rows only") -->

<!-- # gt(head(anzsoc_unit, 10)) %>% -->
<!-- #   tab_header( -->
<!-- #     title = "ANZSOC 2013 - Unit group", -->
<!-- #     subtitle = "first 10 rows only" -->
<!-- #   ) -->


<!-- kable(head(anzsoc_unit, 10),  -->
<!--     caption = "ANZSOC 2013 - Occupation, first 10 rows only") -->

<!-- # gt(head(anzsoc_occupation, 10)) %>% -->
<!-- #   tab_header( -->
<!-- #     title = "ANZSOC 2013 - Occupation", -->
<!-- #     subtitle = "first 10 rows only" -->
<!-- #   ) -->

<!-- ``` -->


## ANZSIC

Merge ANZSIC tables together. `data/processed-data/cleaned_anzsic_2006.csv`

```{r, echo = FALSE}
library(data.table)
library(readxl)
library(magrittr)
library(ggplot2)
library(kableExtra)
# library(gt)

anzsic_codes <- 
  readxl::read_xlsx(path = "data/raw-data/ANZSIC.xlsx", sheet = "codes") %>%
  janitor::clean_names() %>% 
  setDT() %>% 
  .[, anzsic_class_title := trimws(anzsic_class_title)]

anzsic_divisions <- 
  readxl::read_xlsx(path = "data/raw-data/ANZSIC.xlsx", sheet = "divisions") %>%
  janitor::clean_names() %>%
  setDT() %>%
  .[, anzsic_division_title := trimws(anzsic_division_title)]

anzsic_codes <- 
  anzsic_codes %>% 
  .[, anzsic_division := substr(anzsic_code, 1,1)] %>%
  merge(., anzsic_divisions, by = "anzsic_division")

kable(head(anzsic_codes, 10), caption = "ANZSIC 2006 table")
```

<!-- ## Identify essential industries -->

<!-- ```{r} -->
<!-- essential_industries <-  -->
<!--   c("Agriculture, Forestry and Fishing",  -->
<!--     "Electricity, Gas, Water and Waste Services", -->
<!--     "Health Care and Social Assistance", -->
<!--     "Retail Trade", -->
<!--     "Accommodation and Food Services", -->
<!--     "Transport, Postal and Warehousing", -->
<!--     "Public Administration and Safety") -->

<!-- anzsic_codes %>% -->
<!--   .[, is_essential := FALSE] %>% -->
<!--   .[anzsic_division_title %in% essential_industries, is_essential := TRUE] -->
<!-- ``` -->

<!-- ```{r anzsic-plot} -->
<!-- ggplot(anzsic_codes, aes(y = anzsic_division_title, fill = is_essential)) + -->
<!--   geom_bar() + -->
<!--   theme_minimal() -->
<!-- ``` -->


```{r}
# write to processed-data
data.table::fwrite(anzsic_codes, "data/processed-data/cleaned_anzsic_2006.csv")
```


## ANZSOC

Merge ANZSOC tables together. See in `data/processed-data/cleaned_anzsoc_2013.csv`

```{r}
anzsoc <-
  readxl::read_xlsx(path = "data/raw-data/ANZSCO.xlsx", sheet = "Table 5", skip = 5, col_names = FALSE) %>%
  janitor::clean_names() %>%
  setDT(.) %>%
  .[, lapply(.SD, trimws)]

anzsoc_major <-
  anzsoc %>%
  .[!is.na(x1) & !is.na(x2)] %>%
  janitor::remove_empty(., which = "cols") %>%
  setnames(c("major_code", "major_group"))

anzsoc_submajor <-
  anzsoc %>%
  .[!is.na(x2) & !is.na(x3)] %>%
  janitor::remove_empty(., which = "cols") %>%
  setnames(c("submajor_code", "submajor_group"))

anzsoc_minor <-
  anzsoc %>%
  .[!is.na(x3) & !is.na(x4)] %>% 
  janitor::remove_empty(., which = "cols") %>%
  setnames(c("minor_code", "minor_group"))

anzsoc_unit <-
  anzsoc %>%
  .[!is.na(x4) & !is.na(x5)] %>%
  janitor::remove_empty(., which = "cols") %>%
  setnames(c("unit_code", "unit_group"))


anzsoc_occupation <-
  anzsoc %>%
  .[!is.na(x5) & !is.na(x6)] %>% 
  janitor::remove_empty(., which = "cols") %>% 
  setnames(c("occupation_code", "occupation_group", "skill_level"))

anzsoc_codes <-
  copy(anzsoc_occupation) %>%
  .[, `:=`(
    major_code = substr(occupation_code, 1, 1),
    submajor_code = substr(occupation_code, 1, 2),
    minor_code = substr(occupation_code, 1, 3),
    unit_code = substr(occupation_code, 1, 4)
  )] %>%
  merge(anzsoc_major, by = "major_code") %>% 
  merge(anzsoc_submajor, by = "submajor_code") %>% 
  merge(anzsoc_minor, by = "minor_code") %>%
  merge(anzsoc_unit, by = "unit_code") %>%
  setcolorder(c("major_code", "major_group", "submajor_code", "submajor_group", "minor_code", "minor_group", "unit_code", "unit_group")) %>%
  .[, skill_level := gsub(" Aus,.+", "", skill_level)]

kable(head(anzsoc_codes, 10), caption = "ANZSOC 2013 table")
```

## Identify essential workers

```{r}
# essential_occupations <- 
#   c("Delivery Driver",
#     "Postal Delivery Officer",
#     "Social Worker",
#     "Chemist")

essential_minor <- 
  c("Medical Practitioners",
    "Midwifery and Nursing Professionals",
    "Defence Force Members, Fire Fighters and Police",
    "Legal Professionals",
    "Prison and Security Officers",
    "Farmers and Farm Managers",
    "Food Trades Workers",
    "Media Professionals",
    "Child Carers",
    "Electricians",
    "Plumbers",
    "Retail Managers",
    "ICT Managers",
    "Delivery Drivers",
    "Truck Drivers",
    "Automobile, Bus and Rail Drivers",
    "Cleaners and Laundry Workers",
    "Food Process Workers",
    "Packers and Product Assemblers",
    "Farmers and Farm Managers",
    "Prison and Security Officers",
    "Financial and Insurance Clerks",
    "Air and Marine Transport Professionals",
    "Education, Health and Welfare Services Managers",
    "ICT and Telecommunications Technicians",
    "Agricultural, Medical and Science Technicians",
    "Electronics and Telecommunications Trades Workers",
    "Storepersons",
    "Miscellaneous Factory Process Workers",
    "Farm, Forestry and Garden Workers",
    "Freight Handlers and Shelf Fillers",
    "Miscellaneous Labourers",
    "Construction and Mining Labourers",
    "Mobile Plant Operators",
    "Checkout Operators and Office Cashiers",
    "Personal Carers and Assistants",
    "Health and Welfare Support Workers",
    "Education, Health and Welfare Services Managers",
    "Agricultural, Medical and Science Technicians",
    "School Teachers") %>% unique()

# essential_submajor <-
#   c("Hospitality, Retail and Service Managers",
#     "Health Professionals",
#     "Factory Process Workers",
#     "Food Preparation Assistants",
#     "Food Trades Workers")

anzsoc_codes %>%
  .[, is_essential := FALSE] %>% 
  .[minor_group %in% essential_minor, is_essential := TRUE]
```

```{r anzsoc-plot}
ggplot(anzsoc_codes, aes(y = submajor_group, fill = is_essential)) +
  geom_bar() +
  theme_minimal() +
  labs(x = "Number of occupations")
```

```{r}
ggplot(anzsoc_codes, aes(y = skill_level, fill = is_essential)) +
  geom_bar() +
  theme_minimal() +
  labs(x = "Number of occupations", 
       caption = "Highly skilled jobs are ranked 1, unskilled jobs are ranked 5")
```

```{r}
# wrote to processed-data
fwrite(anzsoc_codes, file = "data/processed-data/cleaned_anzsoc_2013.csv")
```


<!-- make plot like these https://github.com/m-jahn/SysbioTreemaps -->

```{r}
occp <- 
  fread("data/raw-data/census-2016-occp.csv") %>%
  janitor::clean_names() %>%
  setDT() %>%
  .[, -c(1)] %>%
  setnames("occp_3_digit_level", "minor_group") %>%
  dcast(minor_group ~ gccsa) %>% 
  janitor::clean_names()

anzsoc_codes <-
  merge(anzsoc_codes, occp, by = "minor_group")


```


```{r cache = TRUE, echo=FALSE}
library(SysbioTreemaps)

tm <- SysbioTreemaps::voronoiTreemap(
  data = anzsoc_codes[is_essential == TRUE, ],
  levels = c("major_group", "minor_group"),
  cell_size = "total",
  shape = "rounded_rect"
)
```


```{r voronoi}
SysbioTreemaps::drawTreemap(tm, label_size = c(2,4), 
            label_level = c(1,2),
            title = "Essential workers")
```

